---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Load the data
```{r}
activity <- read.csv("~/Desktop/Coursera/ReproducibleResearch/RepData_PeerAssessment1/activity.csv",
                     stringsAsFactors=FALSE)
```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day

```{r}
activity_totalsteps <- aggregate(steps ~ date, data = activity, sum)
```

2. Make a histogram of the total number of steps taken each day

```{r fig.width=10, fig.height=8, fig.align='center', fig.cap="Histogram of The Total Number of Steps Taken Each Day"}
    xname <- "Steps Taken Each Day"
    yname <- "Frequency"
    hist(activity_totalsteps$steps,  
         main   = paste("Histogram Of The Total Number Of", xname), 
         xlab   = xname, 
         ylab   = yname, 
         col    = "lightblue", 
         labels = FALSE,
         axes   = "TRUE",
         plot   = "TRUE",
         border = "gray")
```

3. Calculate and report the **mean** and **median** total number of steps taken per day

```{r}
mean(activity_totalsteps$steps)
median(activity_totalsteps$steps)
```
With no NA removed the mean steps per day is `r mean(activity_totalsteps$steps)` and the median steps per day `r median(activity_totalsteps$steps)`

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps_interval <- aggregate(steps ~ interval, data = activity, mean)
plot(steps_interval$interval,
     steps_interval$steps, 
     type="l",
     xlab="Interval",
     ylab="Number of Steps",
     main="The Average Daily Activity Pattern")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxSteps<-steps_interval$interval[which.max(steps_interval$steps)]
```
The interval that contains the maximum number of steps is `r maxSteps`.


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
