---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
###Load ggplot
```{r}
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
```

###Loading and preprocessing the data
1. Download, unzip then Load the data

```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip", method= "curl")
unzip("activity.zip")
activity <- read.csv("activity.csv", header=T)
```

### What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day

```{r}
activity_totalsteps <- aggregate(steps ~ date, data = activity, sum)
```

2. Make a histogram of the total number of steps taken each day

```{r fig.width=10, fig.height=8, fig.align='center', fig.cap="Histogram of The Total Number of Steps Taken Each Day"}
    xname <- "Steps Taken Each Day"
    yname <- "Frequency"
    hist(activity_totalsteps$steps,  
         main   = paste("Histogram Of The Total Number Of", xname), 
         xlab   = xname, 
         ylab   = yname, 
         col    = "lightblue", 
         labels = FALSE,
         axes   = "TRUE",
         plot   = "TRUE",
         border = "gray")
```

3. Calculate and report the **mean** and **median** total number of steps taken per day

```{r}
mean(activity_totalsteps$steps)
median(activity_totalsteps$steps)
```
With no NA removed the mean steps per day is `r mean(activity_totalsteps$steps)` and the median steps per day `r median(activity_totalsteps$steps)`

### What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps_interval <- aggregate(steps ~ interval, data = activity, mean)
steps_interval$time<- seq_along(steps_interval$interval)
plot(steps_interval$time,
     steps_interval$steps, 
     type="l",
     xaxt="n",
     xlab="Time",
     ylab="Number of Steps",
     main="Average Daily Activity Pattern")
axis(1, at = (0:12)*24, labels = paste0(c(0:12)*2, ':00'))
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxSteps<-steps_interval$interval[which.max(steps_interval$steps)]
```
Assuming that the intervals begin at midnight and the intervals represent clock times,the 5-min interval, on average across all the days in the dataset that contains the maximum number of steps is `r maxSteps`.

### Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with ð™½ð™°s)
Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r} 
missing <- sum(is.na(activity$steps))
```
The total number of NA's in the activity dataset was : `r missing`

2. Create a new dataset that is equal to the original dataset but with the missing data filled in.

####find out what rows are missing data and assign it to a vector
```{r}
which_missing  <- which(is.na(activity$steps))
```
####create an index to hold the intervals corresponding to missing data
```{r}
interval_row <- match(activity[which_missing,"interval"], steps_interval$interval)
```
####create a df to impute the data
```{r}
activity_imp <- activity
```
####Impute the missing values to activity_imp using which_missing as the index for the missing values and interval_row as the index for which steps to uses from steps_interval$steps. 
```{r}
activity_imp[which_missing,"steps"] <- steps_interval[interval_row,"steps"]
head(activity_imp)
```

3. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


#### Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

####change activity_imp$date from charachter to date 
```{r}
activity_imp$date<- as.Date(activity_imp$date)
```
#create the new factor variable named `r activity_imp$wday`
```{r}
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')

activity_imp$wday <- factor((weekdays(activity_imp$date) %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
head(activity_imp)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
stepswday_interval <- aggregate(steps ~ interval+wday, data = activity_imp, mean)

stepswday_interval$time <- 1:288

```

```{r}
ggplot(stepswday_interval, aes(time, steps)) + geom_line() + facet_grid(wday ~ .) + 
  scale_x_continuous(breaks = (0:12)*24, labels = paste0(c(0:12)*2, ':00'))

```